### 2014-02-10 text/Abstract.tex
Merge changes
Adding popular introduction, title and new abstract

\chapter*{Abstract} % Abstract name

Static type systems are the most common form of program verification. In their most basic form, type
systems guarantee that a computation always returns a value of some type. The work on monads
and effect systems extends type systems to guarantee properties about \emph{effectful 
computations} that perform side-effects like non-termination, I/O operations or exceptions.

In this thesis we extend type systems to guarantee properties about \emph{context-dependent
computations}. Such computations are increasingly common in modern software. For example, applications
accessing data need to guarantee that the right data is available and distributed systems need to 
guarantee that resources are available on specific nodes.

We develop three \emph{coeffect} type systems that are capable of capturing different context-dependent
properties. All three systems are based on the unifying semantic structure called \emph{indexed
comonads}, but they use it differently.

Our \emph{flat coeffect calculus} tracks one annotation about the entire context. It can be used to
keep a set of required resources or a set of supported platforms. Our \emph{structural coeffect 
calculus} tracks one annotation for each variable. It can be used to track liveness of variables
or number of past values needed in data-flow computations. Finally, our \emph{coeffect meta-language}
embeds contextual values directly into a language. It can be used to guarantee security properties
of computations by using tainting.

>>>>

\chapter*{Abstract} % Abstract name
The development of programming languages needs to reflect important changes in the industry.
In recent years, this included \eg~the development of parallel programming models (in reaction 
to the multi-core revolution). This thesis is a response to another such revolution -- the 
diversification of devices and systems where programs run.

We develop the foundations for statically typed functional languages that understand 
the \emph{context} or environment in which programs execute. Such context includes different 
platforms (and their versions) in cross-platform applications, resources available in different 
execution environments (\eg \; GPS sensor on a phone and database on the server), but also more 
traditional notions such as variable context (tracking variable usage in static analyses) or 
past values in stream-based data-flow programming.

The thesis presents three \emph{coeffect} calculi that capture different notions of 
context-awareness: \emph{flat} calculus capturing contextual properties of the execution 
environment, \emph{structural} calculus capturing contextual properties related to 
variable usage and \emph{meta-language} calculus which allows reasoning about multiple notions
of context-dependence in a single language and provides pathway to embedding the other two
calculi in existing languages.

Although the focus of this thesis is on the syntactical properties of the presented 
systems, we also discuss their category-theoretical motivation. We introduce the notion of
an \emph{indexed} comonad (the dual of monads) and show how they provide semantics of the
presented three calculi.
